---
title: "RWorksheet_Marquez#4c"
author: "Kurt Sayro D. Marquez"
date: "2024-11-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#1a.
```{r}
mpg <- read.csv("/cloud/project/RWorkSheet_Marquez#4a/mpg.csv")
```
#1b.
#The categorical variables are manufacturers, model, trans, drv, fl, and class

#1c.
#The continuous variables are displ, year, cyl, cty, and hwy

#2a. 
```{r}
library(dplyr)

manufacturers <- mpg %>%
  group_by(manufacturer) %>%
  summarise(unique_models = n_distinct(model)) %>%
  arrange(desc(unique_models))
print(manufacturers)

models <- mpg %>%
  group_by(model) %>%
  summarise(variations = n()) %>%
  arrange(desc(variations))


print(models)
```
#2b.
```{r}
barplot(
  manufacturers$unique_models,
  names.arg = manufacturers$manufacturer,
  las = 2,
  col = "purple",
  main = "Number of Unique Models by Manufacturer",
  ylab = "Number of Models"
)
library(ggplot2)


ggplot(manufacturers, aes(x = reorder(manufacturer, -unique_models), y = unique_models)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "Number of Unique Models by Manufacturer", x = "Manufacturer", y = "Number of Models") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#2a.
```{r}
ggplot(mpg, aes(model, manufacturer)) + geom_point()
```
#b.
#For me it is not useful because it lacks context and the text in x-axis is not readable. To make it more readable, I will use barplot to show the count of models for each manufacturer.

```{r}
library(dplyr)
library(ggplot2)

top20 <- mpg %>%
  arrange(desc(year)) %>%
  head(20)

ggplot(top20, aes(x= model, y= year,)) + geom_point() + labs(title = "Top Car Models", xlab="Car model", ylab="Year") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#4a.
```{r}
library(dplyr)
library(ggplot2)

car_counts <- mpg %>%
  group_by(model) %>%
  summarise(number_of_cars = n()) %>%
  arrange(desc(number_of_cars))



top_20_models <- head(car_counts, 20)


ggplot(top_20_models, aes(x = reorder(model, number_of_cars), y = number_of_cars)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#4b.
```{r}
ggplot(top_20_models, aes(x = reorder(model, number_of_cars), y = number_of_cars)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Top 20 Car Models by Number of Cars",
       x = "Car Model",
       y = "Number of Cars") +
  coord_flip() 
```
#5a.Vehicles designed with more cylinders are likely to have larger engine capacities
```{r}
ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +  
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement (litres)") 
```
#6.The result displays the relationship between engine displacement and highway miles per gallon, with points colored based on city mpg 
```{r}
library(ggplot2)


ggplot(mpg, aes(x = displ, y = hwy, color = cty)) +
  geom_point() +  
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement (litres)",
       y = "Highway Miles Per Gallon")
```
#6a.
```{r}
traffic <- read.csv("/cloud/project/RWorkSheet_Marquez#4a/traffic.csv")
obs <- nrow(traffic)
vrbls <- names(traffic)
obs
vrbls
```
#6b.
```{r}

junctions_data <- split(traffic, traffic$Junction)


traffic$DateTime <- as.POSIXct(traffic$DateTime, format="%Y-%m-%d %H:%M:%S")

ggplot(traffic, aes(x = DateTime, y = Vehicles, color = factor(Junction))) +
  geom_line() +
  labs(title = "Traffic Volume Over Time by Junction",
       x = "Date and Time",
       y = "Number of Vehicles",
       color = "Junction") +
  theme_minimal() +
  facet_wrap(~ Junction, scales = "free_y")
  
```
#7a.
```{r}
library(openxlsx)
library(readxl)

alexa <- read.xlsx("/cloud/project/RWorkSheet_Marquez#4a/alexa_file.xlsx")
observation <- nrow(alexa)
columns <- ncol(alexa)
observation
columns
```
#7b.
```{r}
variation_counts <- alexa %>%
  group_by(variation) %>%
  summarise(total = n())
variation_counts
```
#7c. There are only 3 variations that has more counts and 2 lowest counts than the others.
```{r}
ggplot(variation_counts, aes(x = variation, y = total, fill = variation)) +
  geom_bar(stat = "identity") +  
  labs(title = "Total Counts of Variations",
       x = "Variation",
       y = "Total Count") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#7d.
```{r}
library(dplyr)
library(ggplot2)

alexa$date <- as.Date(alexa$date)

daily_reviews <- alexa %>%
  group_by(date) %>%
  summarise(total_verified_reviews = sum(feedback))

ggplot(daily_reviews, aes(x = date, y = total_verified_reviews)) +
  geom_line(color = "green") +
  labs(title = "Number of Verified Reviews Over Time",
       x = "Date",
       y = "Total Verified Reviews") 
```
#7e.
```{r}
library(dplyr)
library(ggplot2)

variation_ratings <- alexa %>%
  group_by(variation) %>%
  summarise(average_rating = mean(rating, na.rm = TRUE)) %>%

arrange(desc(average_rating))

variation_ratings

```
```{r}
ggplot(variation_ratings, aes(x = reorder(variation, -average_rating), y = average_rating, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Rating by Product Variation",
       x = "Product Variation",
       y = "Average Rating") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = "none")
```