---
title: "RWorksheet_Marquez#2"
author: "Kurt Sayro D. Marquez"
date: "2024-09-30"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Create a vector using : operator
#a. The output prints -5 to 5
```{r}
-5:5
```
#b. the value of x is 1-7
```{r}
x <- 1:7
x
```
#2.* Create a vector using seq() function a. seq(1, 3, by=0.2) # specify step size
#the output prints from 1 to 3 with intervals of 0.2

```{r}
seq(1, 3, by = 0.2)
```
#3. A factory has a census of its workers. There are 50 workers in total.
```{r}
age <- list(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41,51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58,26,18)
```
#a. the value is 22
```{r}
age[3]
```
#b. the 2nd value is 28 and the 4th value is 36
```{r}
age[c(2,4)]
```
#c. Access all but the 4th and 12th element is not included. Write the R script and its output.
#the output prints the the age vector without the 4th and 12th element
```{r}
age[-c(4, 12)]
```
#4. *Create a vector x <- c("first"=3, "second"=0, "third"=9). Then named the vector, names(x).
```{r}
names <- c("first"=3, "second"=0, "third"=9)
```
#a. the output prints first, second and third as columns and their values. after entering the code, it only prints the first and the third column.
#b. Write the code and its output.
```{r}
names
names[c("first", "third")]
```
#5. Create a sequence x from -3:2.
```{r}
x <- -3:2
```
#a. Modify 2nd element and change it to 0;
#b. the 2nd element in the output is now 0.
```{r}
x[2] <- 0
x
```
#6. *The following data shows the diesel fuel purchased by Mr. Cruz. a. Create a data frame for month, price per liter (php) and purchase-quantity (liter). 
```{r}
diesel_fuel <- data.frame(
  Month = c("Jan", "Feb", "March", "Apr", "May", "June"),
  Price_per_Liter = c("52.50", "57.25", "60.00", "65.00", "74.25", "54.00"),
  Purchase_Quantity = c("25", "30", "40", "50", "10", "45")
)
diesel_fuel
```
#b. the average fuel expenditure of Mr. Cruz is 59.2625 or 59.26
```{r}
diesel_fuel$Price_per_Liter <- as.numeric(diesel_fuel$Price_per_Liter)
diesel_fuel$Purchase_Quantity <- as.numeric(diesel_fuel$Purchase_Quantity)
avg_fuel_exp <- weighted.mean(diesel_fuel$Price_per_Liter, diesel_fuel$Purchase_Quantity)

avg_fuel_exp
```
#7. R has actually lots of built-in datasets. For example, the rivers data “gives the lengths(in miles) of 141 “major” rivers in North America, as compiled by the US Geological Survey”. a. Type “rivers” in your R console.
```{r}
Na_rivers <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers), sd(rivers), min(rivers), max(rivers))

```
#b.the output prints the length, sum, mean, median, variance, standard deviation, minumum, and maximum.
```{r}
Na_rivers
```
#8. The table below gives the 25 most powerful celebrities and their annual pay as ranked by the editions of Forbes magazine and as listed on the Forbes.com website.
#a. Create vectors according to the above table. Write the R scripts and its output.
```{r}
power_ranking <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25)
celeb_name <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods", "Steven Spielberg", "Howard Stern", "50 Cent", "Cast of the Sopranos", "Dan Brown", "Bruce Sprinsteen", "Donal Trump", "Muhammad Ali", "Paul McCartney", "George Lucas", "Elton John", "David Letterman", "Phil Mickelson", "J.K. Rowling", "Brad Pitt", "Peter Jackson", "Dr. Phil McGraw", "Jay Leno", "Celine Dion", "Kobe Bryant")
pay <-  c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 55, 40, 233, 60, 40, 47, 75, 25, 39, 45, 32, 40, 31)

power_ranking
celeb_name
pay

```
#b.
```{r}
Rowling <- which(celeb_name == "J.K. Rowling")
power_ranking[Rowling] <- 15
pay[Rowling] <- 90

power_ranking
pay

```
#c
```{r}
Forbes <- read.csv("/cloud/project/Worksheet/PowerRanking.csv")

```
#d. 
```{r}
Ranks <- Forbes[10:20,]
Ranks
save(Ranks, file = "/cloud/project/Worksheet/Ranks.RData")
```
#e. the output prints the rows from 10 to 20.
#9. Download the Hotels-Vienna https://tinyurl.com/Hotels-Vienna
#a. Import the excel file into your RStudio. What is the R script?
```{r}
library("openxlsx")
library("readxl")

hotelsvienna <- read_xlsx("/cloud/project/Worksheet/hotels-vienna.xlsx")

```
#b. it has 428 row dimensions and 24 column dimensions
```{r}
dim(hotelsvienna)
```
#c. Select columns country, neighbourhood, price, stars, accomodation_type, and ratings. Write the R script.
```{r}
column <- hotelsvienna[, c("country", "neighbourhood", "price","stars", "accommodation_type", "rating")]
```
#d. Save the data as **new.RData to your RStudio. Write the R script.
```{r}
save(column, file ="/cloud/project/Worksheet/new.RData")
```
#Display the first six rows and last six rows of the new.RData. What is the R script?
```{r}
load("/cloud/project/Worksheet/new.RData")
head(column, 6)
tail(column, 6)
```
#10. a. Write the R scripts and its output.
```{r}
vegetables <- list("Carrot", "Potato", "Onion", "Peas", "Cabbage", "Eggplant", "Broccoli", "Lettuce", "Green Beans", "Garlic")
vegetables
```
#b. the output prints the additional 2 vegetables.
```{r}
vegetables <- append(vegetables, list("Cucumber", "Capsicum"))
vegetables
```
#c. the vegetable list have 16 datapoints.
```{r}
vegetables <- append(vegetables, list("Cauliflower", "Radish", "Tomato", "Beet Root"))
vegetables
length(vegetables)
```
#there are 13 vegetables left.
```{r}
vegetables <- vegetables[-c(5, 10, 15)]
vegetables
```